Sampling Strategies in Decisions from Experience

### 2022-03-30 manuscript.Rmd
Add abstract

abstract: |
  Add abstract

keywords          : ""
wordcount         : ""

>>>>

abstract: "In decisions from experience, systematic choices between prospects are based on the assessment of differences in their latent properties. Yet, distinct properties can be used to distinguish among the prospects, and the choices may vary respectively. This paper presents a model that shows how the mind may rely on sampling strategies---i.e., the probability of sampling outcomes from different prospects in direct succession (switching probability)---to alter which differences in the latent properties evidence is accumulated for. Specifically, the model sequentially samples outcomes from two prospects and accumulates the results obtained from a sequence of ordinal comparisons between sample means, with switching probabilities determining the size of the samples underlying each comparison and thresholds determining how much comparisons a prospect must win to be chosen. In a computational analysis, the model is used to simulate the information-processing chain for choices between two prospects, while varying switching probabilities and thresholds. The analysis shows that the accumulation process provides evidence for differences in the expected values when low switching probabilities cause each comparison to be carried out on large samples. In turn, when high switching probabilities cause the accumulation process to unfold over comparisons based on small samples, evidence is accumulated for differences in the probability to produce a higher outcome most of the time. The respective choices for high switching probabilities reflect an as-if underweighting of rare outcomes that becomes more pronounced with increasing thresholds, demonstrating that the reliance on many small samples can explain the as-if underweighting of rare outcomes, while a large number of sampled outcomes is fully accounted for. The analysis also shows that the differences in the information-processing chain translate to systematic differences in the parameter estimitates of cumulative prospect theory's value and weighting function."

keywords          : "risky choice, information sampling, evidence accumulation, cognitive modeling, cumulative prospect theory"
wordcount         : ""
